### 2018-05-16 paper/paper.Rmd
i cant use epa affiliation unless my advisors are on the project .... easier to just do this :)

---
title: S-Tipping vs B-Tipping
## We can figure out author order later on who end up doing more etc :-)
author:
  - name: Ryan Batt
    affiliation: Rensselaer Polytechnic Institute and Rutgers University
  - name: Carl Boettiger
    email: cboettig@berkeley.edu
    affiliation: espm
    footnote: Corresponding Author

address:
  - code: espm
    address: Department of Environmental Science, Policy, and Management, University of California, Berkeley, CA 94720
  - code: epa
    address: Environmental Protection Agency, ...
abstract: |
  This is the abstract.

layout:  3p # review = doublespace, 3p = singlespace, 5p = two-column
preamble: |
  \newcommand{\ud}{\mathrm{d}}
  \usepackage[nomarkers]{endfloat}
  \linenumbers
  \usepackage{setspace}
  \doublespacing
journal: Journal of Mathematical Biology
bibliography: refs.bib
csl: journal-of-mathematical-biology.csl
output: rticles::elsevier_article

keywords: tipping point, bifurcation, saddle-node, state
---

# Introduction

Critical transistions.

Two flavors: S tipping vs B tipping

Dimensionality.  Absense of an explicit model often makes it difficult to intuit if and how a complex ecosystem can be intepreted in terms of this simple one-dimensional bifurcation process.  Lack of clarity about what defines the "state" and what defines the "parameters" leads to confusion.  These issues are easily glossed over...

Focus on a prominent example from lake ecosystems.

# Model

@Carpenter2008

# Analysis

Variable | Description
-------- | ----------------
 $A$     | Adult Bass
 $J$     | Juvenile Bass
 $F$     | Planktivore zooplankton
 $H$     | Herbivorous zooplankton
 $P$     | Phytoplankton

The original model presented in @Carpenter2008 is defined in terms of continuous dynamics within years coupled to discrete annual dynamics.  Within a given year:

$$\frac{d A}{dt} = - q E A$$
$$d F = \left(D_f (F_R - F) - c_{FA} FA \right) dt  + \sigma_F dW_f$$

$$ \frac{dJ}{dt} = -c_{JA} JA - \frac{c_{JF} v FJ}{h + v + c_{JF} F}$$

While dynamics between years are defined in terms of integrals over the within-year dynamics:

$$A_{t+1} = s \int_t^{t+i} \left(A_{t} + J_{t} \right) dt $$
$$F_{t+1} = \int_t^{t+i} F_{t} dt $$

$$J_{t+1} = fsA_{t+1}$$

Based on these equations, we can approximate these dynamics in continuous time as:

$$\frac{d A}{dt} = - q E A_t + sJ $$
$$d F = \left(D_f (F_R - F) - c_{FA} FA \right) dt  + \sigma_F dW_f$$

$$\frac{dJ}{dt} = -c_{JA} JA - \frac{c_{JF} v FJ}{h + v + c_{JF} F} + fA$$

Dynamics for herbivorous plankton and phytoplankton are given only in continuous time:

$$dH = \left(D_H (H_R - H) + \alpha c_{HP} HP - c_{HF} HF\right) dt + \sigma_H dW_h$$

$$dP = \left( r_P L \gamma(I_o, P) P  - m P - c_{PH} HP \right) dt + \sigma_P dW_P$$

### Non-Dimensionalization

$$\tilde F := D_f (F - F_R) $$
$$\frac{d A}{dt} = - q E A_t + sJ $$
$$d \tilde F = - \left(1 + c_{fA} A \right) \tilde F  dt  + \sigma_F dW_f$$
$$\beta := \frac{h+v}{c_{JF}} $$

$$\frac{d J}{dt} = fA -\left(c_{JA} A + \frac{v F}{\beta + F} \right) J $$

$$\tilde H := D_H (H - H_R) $$
$$d \tilde H = \tilde H \left(1 + \alpha c_{\tilde HP} P - c_{\tilde HF} F\right) dt + \sigma_H dW_h$$

### Dimensionality reduction

## Implications for Resilience and Early Warning

# Discussion

# Acknowledgements

CB acknowledges support in part from USDA National Institute of Food and Agriculture, Hatch project CA-B-INS-0162-H.

\newpage

# References