### 2018-11-09 analysis/manuscript_report.Rmd
R -> Rmd in analyses

---
title: "Bifurcation or state tipping: assessing transition type in a model trophic cascade"
author: "Ryan Batt & Carl Boettiger"
date: "2018-07-07"
abstract: |
       The primary purpose of this document is to sketch out the key equations, figures, and numerical results for a manuscript that focuses on assessing whether a model undergoes state-tipping or a bifurcation. Our approach emphasizes reducing the dimensionality of the original model to redefine manipulated state variables as parameters. Thus, when these redefined parameters are manipulated to induce a transition, because they parameters not states, the transition should be viewed as a bifurcation, not state tipping.
output:
  # html_document:
  #   toc: true
  #   toc_depth: 4
  #   fig_caption: true
  #   theme: "readable"
  #   template: default
  pdf_document:
    toc: true
    toc_depth: 4
    keep_tex: true
#    template: latex-ryan.template
    fig_caption: yes
geometry: margin=1.0in
lineno: true
lineSpacing: false
titlesec: true
documentclass: article
placeins: true
---

```{r report-setup, include=FALSE, echo=FALSE, cache=FALSE}
# ==========================================
# = Record Time to Get Elapsed Time at End =
# ==========================================
t1 <- Sys.time()

# =================
# = Load Packages =
# =================
library(viridis)
library(phaseR)
library(rootSolve)
library(bs.tipping)

# Report
library(knitr)
library(rmarkdown)

# ================
# = Report Setup =
# ================
doc_type <- c("html", "pdf")[1]
table_type <- c("html" = "html", "pdf" = "latex")[doc_type]
options("digits" = 3) # rounding output to 4 in kable() (non-regression tables)
o_f <- paste(doc_type, "document", sep = "_")

opts_chunk$set(
  fig.path = "manuscript_report/",
  cache.path = "manuscript_report/",
  echo = TRUE,
  include = TRUE,
  cache = FALSE,
  autodep = TRUE,
  results = "asis",
  warning = FALSE,
  fig.show = "hold",
  fig.lp = NULL
)
```

# System of Equations & Manipulation

Start with the equations for the ecosystem dynamics:

$$\begin{array}{llr}
\dot{A} &= sJ - qEA - m_AA &(1) \\
\dot{F} &= D_F(F_o - F) - c_{FA}FA &(2)\\
\dot{J} &= fA - c_{JA}JA - \frac{c_{JF}vJF}{h+v+c_{JF}F} - sJ &(3)\\
\dot{H} &= D_H(H_R-H) + \alpha c_{HP}HP - c_{HF}HF &(4)\\
\dot{P} &= r_P L \gamma (I_0,P)P - mP - c_{PH}HP &(5)\\
\end{array}$$

Focusing on the fish, and setting each of the 3 fish differential equations to 0 and solving for the respective state variables:

$$\begin{array}{llr}
A&= \frac{sJ}{qE+m_A} &(6)\\[10pt]
F&=\frac{D_FF_0}{D_F + c_{FA}A} &(7)\\[10pt]
J&=\frac{fA}{c_{JA}A+s+\frac{c_{JF}vF}{h+v+c_{JF}F}} &(8)\\[10pt]
\end{array}$$

**Begin Equations of Motion**  
Although our ultimate goal is to see if the equation of motion for $J$ or $F$ shows alternate states, for completeness we will also present the equation of motion for $A$, which is acquired by substituting the RHS of Eq7 for $F$ in Eq8, then substituting the resulting expression for $J$ in Eq1:

$$\begin{array}{llr}
\dot{A} &=  sfA\Bigg(c_{JA}A+s+c_{JF}v\bigg(\frac{(h+v)(D_F+c_{FA}A)}{D_F F_o}+c_{JF}\bigg)^{-1}\Bigg)^{-1} -qA -m_A A &(9)
\end{array}$$

To get the equation of motion for $F$, we begin by getting an expression for $A$ that depends only on $F$ and parameters, which is achieved by substituting Eq8 into Eq6 and solving for $A$:

$$\begin{array}{llr}
A&=\bigg({\frac{sfA}{c_{JA}A+s+\frac{c_{JF}vF}{h+v+c_{JF}F}}}\bigg)\bigg(qE+m_A\bigg)^{-1} &(10a)\\[15pt]
sfA &= A(qE+m_A)\big(c_{JA}A+s+\frac{c_{JF}vF}{h+v+c_{JF}F}\big) &(10b)\\[10pt]
sf &=(qE+m_A)\big(c_{JA}A+s+\frac{c_{JF}vF}{h+v+c_{JF}F}\big) &(10c)\\[10pt]
A&=\frac{sf}{c_{JA}(qE+m_A)} - \frac{s}{c_{JA}} - \frac{c_{JF}vF}{c_{JA}(h+v+c_{JF}F)} &(10d)
\end{array}$$

Going from Eq10b to Eq10c is achieved by dividing both sides of the equation by $A$, which implies that in Eq10c $A$ cannot be 0 when $\dot{A}$ is 0. We will return to the consequences of this assumption later as we assess our derived equation of motion for $F$. By subtituting the RHS of Eq10d for $A$ in Eq2, we arrive at the equation of motion for $F$:

$$\begin{array}{llr}
\dot{F} &= D_F(F_o-F) - c_{FA}F(\frac{sf}{c_{JA}(qE+m_A)} - \frac{s}{c_{JA}} - \frac{c_{JF}vF}{c_{JA}(h+v+c_{JF}F)}) &(11)
\end{array}$$

Eq10 is what is used in `dF_dt_1state`.

The equation of motion for $J$ can be acquired by substituting the RHS of Eq7 for $F$ in Eq3 ,then exchanging $A$ in that resulting expression for the RHS of Eq6:

$$\begin{array}{llr}
Q &\equiv 1/(qE + m_A) \\
 \dot{J} &= fsJQ - c_{JA}sQJ^2 - sJ - \frac{c_{JF}vJ}{\frac{(h+v)}{(D_FF_o)(D_F+c_{FA}sQJ)^{-1}} + c_{JF}} &(12)
\end{array}$$

***

# Table of Parameter Values

```{r parameterTable, results='markup', echo=FALSE}

# Symbol
# Description
# Quantity
# Units
paramDefs <- c(
  # "A",
  # "Adult bass",
  # "variable",
  # "kg ha^{-1}", # so, it's actually number of fish, but if you multiply by 0.2 kg/fish you get kg/ha
  #
  # "F",
  # "Planktivorous fish",
  # "variable",
  # "kg ha^{-1}", # actually n, but divide by 500 to get kg/fish
  #
  # "J",
  # "Juvenile bass",
  # "variable",
  # "kg ha^{-1}", # actually n, but divide by 10 to get kg/ha
  #
  # "H",
  # "Herbivorous zooplankton",
  # "variable",
  # "", # no biomass conversion listed in carpenter et al 2008 table
  #
  # "P",
  # "Phytoplankton",
  # "variable",
  # "", # no biomass conversion listed in carpenter et al 2008 table

  "$q$",
  "Harvest or stocking",
  "variable",
  "$t^{-1}$",

  "$s$",
  "Survival rate of juveniles to maturation",
  "0.6",
  "$t^{-1}$",

  "$m_A$",
  "Mortality rate of adults",
  "0.4",
  "$t^{-1}$",

  "$m$",
  "Phytoplankton mortality",
  "0.1",
  "$t^{-1}$",

  "$f$",
  "Fecundity of adult bass",
  "2",
  "$J/A$",

  "$c_{FA}$",
  "Predation of planktivores by adult bass",
  "0.3",
  "$t^{-1} A^{-1}$",

  "$c_{JA}$",
  "Predation of juvenile bass on by adult bass",
  "0.1",
  "$t^{-1} A^{-1}$",

  "$c_{JF}$",
  "Predation of juvenile bass by planktivorous fish",
  "0.5",
  "$t^{-1} F^{-1}$",

  "$c_{PH}$",
  "Predation of phytoplankton by herbivores",
  "0.25",
  "$t^{-1} H^{-1}$",

  "$c_{HF}$",
  "Predation of herbivores by planktivores",
  "0.1",
  "$t^{-1} F^{-1}$",

  "$F_o$",
  "Abundance of planktivorous fish in non-foraging arena",
  "200",
  "F",

  "$H_R$",
  "Reservoir of herbivores in a deep-water refuge",
  "4",
  "H",

  "$D_F$",
  "Diffusion of planktivores between refuge and foraging arena",
  "0.09",
  "$t^{-1}$",

  "$DH$",
  "Diffusion of between shallow and deep water",
  "0.5",
  "$t^{-1}$",

  "$v$",
  "Rate at which J enter a foraging arena, becoming vulnerable",
  "80",
  "$t^{-1}$",

  "$h$",
  "Rate at which J hide in a refuge",
  "80",
  "$t^{-1}$",

  "$r_P$",
  "Growth rate of phytoplankton",
  "3",
  "$m^2 mg^{-1}$",

  "$L$",
  "Phosphorus loading rate",
  "0.6",
  "$mg m^{-2} t^{-1}$",

  "$\\alpha$",
  "Assimilation of phytoplankton phosphorus by zooplankton",
  "0.3",
  "$g g^{-1}$",

  "$\\gamma$",
  "Density-dependent phytoplankton growth response to light",
  "n/a",
  "n/a",

  "$I0$",
  "Solar irradiance incident on the surface of the water",
  "300",
  "$\\mu E m^{-2} s^{-1}$"
)
colNames <- c("Symbol", "Description", "Quantity", "Units")
paramTable_mat <- matrix(paramDefs, byrow = TRUE, ncol = 4, dimnames = list(NULL, colNames))
kable(paramTable_mat, caption = "**Table 1.** Model parameters, description, values, and units. Parameters are taken from Carpenter et al (2008) and Carpenter et al (2013). The function $\\gamma$ describes phytoplankton growth, and can be found in Carpenter et al (2008).")
```

# Setup

```{r setup}
set.seed(42)
qE <- -0.001 # 0.02
qE_end <- 0.001 # /20 #-5#-0.025

dt <- 0.01 # 0.01
nYears <- 1020 # /20

noise_coeff <- c(0.01, 0.01, 0.01, 0.01, 0.01)
```

# Simulation
## Simulate Time series

```{r simulation}
# set initial values to equilibrium (roots)
X_init <- pmax(bs.tipping::getRoot5D(c(A0 = 150, F0 = 50, J0 = 100, H0 = 100, P0 = 500),
                                     pars = c(qE = qE)), 1E-3)

# simulate an example of fish
qE_vec <- seq(qE, qE_end, length.out = nYears / dt)
stateMat <- matrix(NA, nrow = nYears / dt, ncol = 5, 
                   dimnames = list(NULL, c("A0", "F0", "J0", "H0", "P0")))
stateMat[1, ] <- unlist(X_init)

for (j in 2:nrow(stateMat)) { # iterate through time steps
  state <- stateMat[j - 1, ]
  dState_dt <- ecoStep(X = state, pars = c(qE = (qE_vec[j])))

  # Euler Method Approximation
  dState <- dState_dt * dt + rnorm(5, sd = c(noise_coeff)) * dt
  eulerState <- pmax(state + dState, 1E-3)
  stateMat[j, ] <- eulerState # euler
  # stateMat[j,] <- getRoot5D(stateMat[j-1,], pars=c(qE=qE_vec[j]))
}

tsMat <- cbind(time = seq(0, nYears - dt, by = dt), qE = qE_vec, stateMat)
```

## Figure: 5-dimension simulation

```{r figure-5dsimulation, fig.width=3.5, fig.height=6, fig.cap="Simulated time series of adult bass (A0), planktivores (F0), and juvenile bass (J0) over a gradient of harvesting of adult bass (qE)."}
par(mfrow = c(5, 1), mar = c(1.5, 2, 0.75, 0.25), ps = 8, 
    cex = 1, mgp = c(1, 0.25, 0), tcl = -0.15)
stateLabs <- c(
  A0 = "A (Adult Bass)",
  F0 = "F (Planktivores)",
  J0 = "J (Juvenile Bass)",
  H0 = "H ( Herbivores)",
  P0 = "P (Phytoplankton)"
)
stateNames <- names(stateLabs)
for (i in 1:length(stateNames)) {
  plot(tsMat[, "time"], tsMat[, stateNames[i]], type = "l",
       ylab = stateLabs[stateNames[i]], xlab = "")
}
```

***

# Equations of Motion
## Figure: A Motion

```{r figure-eqMotion-A, fig.width=3.5, fig.height=3.5, fig.cap="Equation of motion for A."}
qevals <- c(0.25, 0.025, 0, -0.025)
par(mfrow = c(2, 2), mar = c(1.85, 1.85, 0.75, 0.25),
    mgp = c(1, 0.15, 0), tcl = -0.15, ps = 9, cex = 1)
for (i in 1:length(qevals)) {
  curve(stateMotion(state0 = x, pars = c(qE = qevals[i])),
        from = 0, to = 20, ylab = "dA/dt", xlab = "A")
  abline(h = 0, lty = 2)
  mtext(paste0("qE=", round(qevals[i], 2)),
        side = 3, line = 0, adj = 0.05, font = 2)
}
```

Looks good, the second stable node appears when qE>0

***

## Figure: J Motion

```{r figure-eqMotion-J, fig.width=3.5, fig.height=3.5, fig.cap="Equation of motion for J."}
par(mfrow = c(2, 2), mar = c(1.85, 1.85, 0.75, 0.25),
    mgp = c(1, 0.15, 0), tcl = -0.15, ps = 9, cex = 1)
for (i in 1:length(qevals)) {
  curve(stateMotion(state0 = x, stateVar = "J0", 
                    pars = c(qE = qevals[i])), 
        from = 0, to = 20, ylab = "dJ/dt", xlab = "J")
  abline(h = 0, lty = 2)
  mtext(paste0("qE=", round(qevals[i], 2)), 
        side = 3, line = 0, adj = 0.05, font = 2)
}
```

As with the adults, the second stable node appears when qE>0

***

## Figure: F Motion

```{r figure-eqMotion-F, fig.width=3.5, fig.height=3.5, fig.cap="Equation of motion for F."}
par(mfrow = c(2, 2), mar = c(1.85, 1.85, 0.75, 0.25),
    mgp = c(1, 0.15, 0), tcl = -0.15, ps = 9, cex = 1)
for (i in 1:length(qevals)) {
  curve(stateMotion(state0 = x, stateVar = "F0", pars = c(qE = qevals[i])),
        from = 0, to = 20, ylab = "dF/dt", xlab = "F")
  abline(h = 0, lty = 2)
  mtext(paste0("qE=", round(qevals[i], 2)), 
        side = 3, line = 0, adj = 0.05, font = 2)
}
```

Unfortunately, this state variable only shows a trivial stable node and a saddle point. This is the state variable that requires algebra cancelling the 'A' state variable in order to get its equation of motion (without A canceling, you cannot get a dF/dt vs F type plot easily [as a function of F and parameters]). We think this result is invalid, as there must be some assumptions in the algebra that invalidate our desired interpretation.

***

## Figure: Combined A, J, F Motion

```{r figure-eqMotion-AJF, fig.width=5, fig.height=5, fig.cap="Motion of A, J, and F. Note that the equation of motion for F failed to exhibit alternate states. In formulating the equation for motion of F, there is an algebraic step that requires dropping from terms via division by A, which assume that $A!=0$. Therefore, states of F that correspond to $A=0$ are invalid. The equation of motion for A indicates that one of the stable nodes is at $A=0$, implying that the equation of motion for F is inaccurate in the corresponding region."}
xlabs <- c("A", "J", "F") # ylabs <- c("dA/dt", "dJ/dt", "dF/dt")
ylabs <- paste0("d", xlabs, "/dt")
svars <- paste0(xlabs, "0")
par(mfrow = c(3, 3), mar = c(1.85, 1.85, 0.75, 0.25), 
    mgp = c(1, 0.15, 0), tcl = -0.15, ps = 9, cex = 1)
for (s in 1:3) {
  for (i in 2:length(qevals)) {
    curve(stateMotion(state0 = x, stateVar = svars[s], pars = c(qE = qevals[i])),
          from = 0, to = 20, ylab = ylabs[s], xlab = xlabs[s])
    abline(h = 0, lty = 2)
    mtext(paste0("qE=", round(qevals[i], 2)), side = 3, line = 0, adj = 0.05, font = 2)
  }
}
```

***

## Figure: A Motion A Ball-in-Cup

```{r}
Adot <- function(A, pars = c(qE = -0.001)) {
  parsF <- unlist(
    formals(ecoStep)[c("fA", "cJA", "cJF", "cFA", 
                       "vuln", "hide", "surv", "Fo", "DF")])
  if (missing(pars)) { 
    # if a function requires qE, pars needs to be supplied (e.g., pars=c(qE=0.1))
    pars <- parsF
  } else {
    pars <- c(pars, parsF[!names(parsF) %in% names(pars)])
  }
  with(as.list(pars), {
    surv * fA * A * (cJA * A + surv + cJF * vuln * 
                       (((hide + vuln) * (DF + cFA * A)) / 
                          (DF * Fo) + cJF)^(-1))^(-1) - 
      qE * A - (mA <- 0.4) * A
  })
}
```

```{r include = FALSE, eval=FALSE}
# Avec <- seq(0, 20, length.out=200)
# dAdt_A_vec <- Adot(A=Avec, pars=c(qE=-0.001))
# plot(Avec, dAdt_A_vec, type='l')
# # curve(Adot(A=x, pars=c(qE=-0.001)), from=0, to=20); abline(h=0)
# plot(Avec, cumsum(dAdt_A_vec), type='l')

# Avec <- seq(0, 20, length.out=200)
# qevals2 <- c(qevals[-1], -1)
# par(mfcol=c(2,4), mar=c(1.85,1.85,0.75,0.25), mgp=c(1,0.15,0), tcl=-0.15, ps=9, cex=1)
# l <- 0
# for(q in 1:length(qevals2)){
# 	dAdt_A_vec <- Adot(A=Avec, pars=c(qE=qevals2[q]))
# 	par(mar=c(1.85,1.85,0.75,0.25))
# 	plot(Avec, dAdt_A_vec, type='l', ylab="dA/dt", xlab="A")
# 	qlab <- paste0("q=",round(qevals2[q],2))
# 	mtext(qlab, side=3, line=0, adj=0.05, font=2)
# 	abline(h=0, lty=2)
# 	mtext(LETTERS[l<-l+1], side=3, line=-1, adj=0.05)
#
# 	par(mar=c(1.0,1.85,1.6,0.25))
# 	plot(Avec, -cumsum(dAdt_A_vec), type='l', ylab="", xlab="", xaxt='n', yaxt='n')
# 	mtext(qlab, side=3, line=0, adj=0.05, font=2)
# 	mtext("A", side=1, line=0)
# 	mtext(LETTERS[l<-l+1], side=3, line=-1, adj=0.05)
# }
```

```{r figure-Amotion-ballincup, fig.width=6, fig.height=3.5, fig.cap="Motion of A, and its integral (ball-in-cup diagram)."}
Avec <- seq(-0.75, 20, length.out = 200)
qevals2 <- c(qevals[-1], -0.5)
par(mfrow = c(2, 4), mar = c(1.85, 1.85, 0.75, 0.25), 
    mgp = c(1, 0.15, 0), tcl = -0.15, ps = 9, cex = 1)
l <- 0
for (q in 1:length(qevals2)) {
  dAdt_A_vec <- Adot(A = Avec, pars = c(qE = qevals2[q]))
  qlab <- paste0("q=", round(qevals2[q], 2))
  # par(mar=c(1.85,1.85,0.75,0.25))
  ylim <- range(dAdt_A_vec)
  if (ylim[2] < 0.5) {
    yadj <- diff(ylim) * 0.15
    ylim[2] <- ylim[2] + yadj # max(ylim[2], 0.15)
  }
  plot(Avec, dAdt_A_vec, type = "l", ylab = "dA/dt", xlab = "A", ylim = ylim)
  mtext(qlab, side = 3, line = 0, adj = 0.05, font = 2)
  abline(h = 0, lty = 2)
  mtext(LETTERS[l <- l + 1], side = 3, line = -0.8, adj = 0.025)
}
for (q in 1:length(qevals2)) {
  dAdt_A_vec <- Adot(A = Avec, pars = c(qE = qevals2[q]))
  qlab <- paste0("q=", round(qevals2[q], 2))
  # par(mar=c(1.0,1.85,1.6,0.25))
  plot(Avec, -cumsum(dAdt_A_vec), type = "l", 
       ylab = "", xlab = "A", xaxt = "s", yaxt = "n")
  mtext(qlab, side = 3, line = 0, adj = 0.05, font = 2)
  # mtext("A", side=1, line=0)
  mtext(LETTERS[l <- l + 1], side = 3, line = -1, adj = 0.025)
}
```

***

# Figure: Bifurcation Diagram

```{r }
qe_range <- c(0.075, -0.1) # range of q values
qevec_bd <- do.call(seq, c(as.list(qe_range), length = 2E3)) # sequence of q values
rts <- vector("list", length = length(qevec_bd)) # empty list to store roots for each q value
for (q in 1:length(qevec_bd)) { # for each val of q
  tq <- qevec_bd[q] # temp q value
  trts <- rootSolve::uniroot.all(f = Adot, interval = c(0, 20),
                                 pars = c(qE = tq)) # temp roots
  rts[[q]] <- trts # store temp root
  tclasses <- vector("character", length = length(trts))
  for (r in 1:length(trts)) { # for each root found for this val of q
    trt <- trts[r] # temp root
    t_perturb <- max(0.01 * trt, 0.01)
    resp_neg <- Adot(trt - t_perturb, pars = c(qE = tq))
    resp_pos <- Adot(trt + t_perturb, pars = c(qE = tq))
    if (resp_neg > 0 & resp_pos < 0) { # crosses 0 line downward
      tclass <- "snode"
    } else if (sign(resp_neg) == sign(resp_pos)) { # barely touches 0 line
      tclass <- "unode"
    } else if (resp_neg < 0 & resp_pos > 0) { 
      # cross 0 line upward; 
      tclass <- "unode"
    }
    tclasses[r] <- tclass
  }
  names(rts[[q]]) <- tclasses
}
```

```{r}
# names(rts) <- paste0('q',round(qevec_bd,4))
get_bb <- function(x) min(x[names(x) %in% "snode"])
get_ub <- function(x) {
  snode_ind <- names(x) %in% "snode"
  if (sum(snode_ind) > 1) {
    return(unname(max(x[snode_ind])))
  } else {
    return(NA)
  }
}
```

```{r}
get_separatrix <- function(x) {
  unode_ind <- names(x) %in% "unode"
  if (any(unode_ind)) {
    return(unname(x[unode_ind]))
  } else {
    return(NA)
  }
}
```

```{r figure-bifurcationDiagram, fig.width=3.5, fig.height=3.5, fig.cap="Bifurcation diagram. Solid line is a stable state. Dashed line is an unstable state."}
bot_branch <- sapply(rts, get_bb)
upper_branch <- sapply(rts, get_ub)
separatrix <- sapply(rts, get_separatrix)
ylim <- range(c(bot_branch, upper_branch, separatrix), na.rm = TRUE)
# xlim <- rev(range(qevec_bd))
# plot(qevec_bd, bot_branch, ylim=ylim, xlim=xlim, type='l')
par(mar = c(2, 1.85, 0.75, 0.25), mgp = c(1, 0.15, 0), tcl = -0.15, ps = 9, cex = 1)
plot(qevec_bd, bot_branch, ylim = ylim, type = "l", 
     ylab = "A (adult bass)", xlab = "q (stocking or harvest)")
lines(qevec_bd, upper_branch, lty = 1)
lines(qevec_bd, separatrix, lty = 2)
```

***

# Session Info

```{r sessionInfo, results='markup'}
difftime(Sys.time(), t1) # how long it took to run these models/ produce this report
Sys.time()
sessionInfo()
```